Viscous Erosion in Porous Media

### 2018-02-28 docs/methods/erosion_methods.tex
NM
Working on figure for 6 body simulation.

We therefore impose the condition of a fixed pressure difference by controlling the incoming flow-rate $\umax$ at each time-step. This is achieved by first computing the Stokes flow with $\umax = 1$ prescribed, then measuring the pressure at a location upstream ($x = -2$) and downstream ($x=2$) of the eroding bodies. These upstream and downstream pressures are given by
\begin{align}
& p_{-} = \frac{1}{2} \int_{-1}^{1} p(-2,y) \, dy \, , \\
& p_{+} = \frac{1}{2} \int_{-1}^{1} p(+2,y) \, dy \, .
\end{align}
We then rescale the computed density function by the pressure difference
\begin{align}
\eeta \leftarrow \frac{\eeta}{(p_{-} - p_{+})} \, ,
\end{align}
This rescaling maintains a {\em fixed} pressure difference. Because the Stokes equations are linear, such a rescaling is permissible and does not alter the accuracy of our method. Furthermore, rescaling by a different factor at each time-step is also permissible because the Stokes equations are memoryless (i.e.~there is no time derivative present in the PDEs).

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.9 \textwidth]{./figs/06bod.pdf}
\caption{\label{fig:50bodies} 6 bodies
}
%A time lapse of 50 bodies eroding in a viscous flow.  The boundary conditions are a constant pressure drop that creates a flow from the left to right.  The rate of erosion is proportional to the shear stress which is equivalent to the vorticity evaluated on the bodies.  The colors represent the vorticity, so regions with large vorticity (in magnitude) are regions where the flow will erode the fastest.
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1

@@

>>>>

We therefore impose the condition of a fixed pressure difference by controlling the incoming flow-rate $\umax$ at each time-step. This is achieved by first computing the Stokes flow with $\umax = 1$ prescribed, then measuring the pressure at a location upstream ($x = -2$) and downstream ($x=2$) of the eroding bodies. These upstream and downstream pressures are given by
\begin{align}
p_{-} = \frac{1}{2} \int_{-1}^{1} p(-2,y) \, dy \, , \qquad
p_{+} = \frac{1}{2} \int_{-1}^{1} p(+2,y) \, dy \, .
\end{align}
We then rescale the computed density function by the pressure difference
\begin{align}
\eeta \leftarrow \frac{\eeta}{(p_{-} - p_{+})} \, .
\end{align}
This rescaling maintains a {\em fixed} pressure difference. Because the Stokes equations are linear, such a rescaling is permissible and does not alter the accuracy of our method. Furthermore, rescaling by a different factor at each time-step is also permissible because the Stokes equations are memoryless (i.e.~there is no time derivative present in the PDEs).

@@

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Convergence test for multiple-body erosion with fixed pressure drop.
}
\vspace{0.3 pc}
\label{convtab}
\begin{tabular}{l l l l}
\hline
\hspace{0.0pc} $\Delta t/t_s$
\hspace{0.5pc} & error 
\hspace{0.5pc} & order
\hspace{0.5pc} & CPU time \\
\hline
%
1/10     	& 2.67E-5  	& --        	& 14 mins  	\\
1/20     	& 7.23E-6  	& 1.88 	& 27 mins  	\\
1/40     	& 1.95E-6  	& 1.89 	& 52 mins  	\\
1/80     	& 4.90E-7  	& 1.99 	& 1.7 hours	\\
1/160     	& 1.25E-7  	& 1.97  	& 3.4 hours	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Parameters in this test are: 20circ128, nits = 5, dt0 = 4e-4, tfin = 4e-3
% epsfac = 10, sigfac = 10, fixarea = 0, fixprdop = 1

@@

6 body simulations. Point out that the bodies tend to develop flat faces wherever there is a nearby body (now that the bodies are numbered, can refer to specific bodies). Point out high vorticity in this channels between bodies. Spontaneous channelization. Point out that our algorithm allows bodies to vanish seamlessly, for example body 4.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.9 \textwidth]{./figs/06bod.pdf}
\caption{\label{fig:50bodies} 6 bodies. TO DO HERE: Number the bodies so that I can describe them.
}
%A time lapse of 50 bodies eroding in a viscous flow.  The boundary conditions are a constant pressure drop that creates a flow from the left to right.  The rate of erosion is proportional to the shear stress which is equivalent to the vorticity evaluated on the bodies.  The colors represent the vorticity, so regions with large vorticity (in magnitude) are regions where the flow will erode the fastest.
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1
